% !TEX root = paper.tex
\documentclass[paper.tex]{subfiles}

\begin{document}

\section{Demonstration of use cases}
\label{sec:examples}

Here we present three examples of axioms being tested to illustrate how the algorithms are used and how they work.
We use an ontology $O$ which consists of the following axioms.
\begin{gather*}
  \oaxiom{DisjointClasses}(\oent{Animal}, \oent{Plant}) \\
  \begin{aligned}
    \oent{Giraffe} &\sqsubseteq \oent{Mammal} \\
    \oent{Mammal} &\sqsubseteq \oent{Animal} \\
    \oent{Herbivore} &\equiv \forall \oent{eats} .\, \oent{Plant} \\
    \oent{Susan} &\aligncolon \oent{Giraffe}
  \end{aligned}
\end{gather*}

Example \ref{exmp:testSubClassSimple} is straightforward and falls into the use case of testing something we expect to be entailed to assure the quality of the ontology.

\begin{example}
  \label{exmp:testSubClassSimple}
  Test that $\oent{Giraffe}$ is a subclass of $\oent{Animal}$.
  In terms of our formal model, this means finding the result of
  \[ \test_O(\oent{Giraffe} \sqsubseteq \oent{Animal}) \]
  In OWL 2 functional syntax, the axiom is represented as
  \[ \oaxiom{SubClassOf}(\oent{Giraffe}, \oent{Animal}) \]
  So we call algorithm \ref{alg:testSubClassOf}.
  \[ \Call{testSubClassOf}{\oent{Giraffe}, \oent{Animal}} \]
  The algorithm first checks if there are any instances of the class expression $\oent{Giraffe} \sqcap \neg \oent{Animal}$ at line \ref{alg:testSubClassOf:checkInconsistent}.
  There are none in this ontology, so it proceeds to check at line if the same class expression has any named subclasses or equivalent classes \ref{alg:testSubClassOf:checkIncoherent}.
  Again there are none, so it checks if the class expression is satisfiable at line \ref{alg:testSubClassOf:checkAbsent}.
  It is not, so the algorithm returns entailed.
\end{example}

Next, examples \ref{exmp:testSubClassComplex} and \ref{exmp:testClassAssertion} demonstrate testing of axioms which are not entailed.
This could arise from mistakes in the ontology or in a test, or from temporary tests used for investigation.

\begin{example}
  \label{exmp:testSubClassComplex}
  Test that $\exists \oent{eats} .\, \oent{Animal} \sqsubseteq \oent{Herbivore}$.
  As in example \ref{exmp:testSubClassSimple}, we call algorithm \ref{alg:testSubClassOf}.
  \[ \Call{testSubClassOf}{\exists \oent{eats} .\, \oent{Animal}, \oent{Herbivore}} \]
  This checks if $\exists \oent{eats} .\, \oent{Animal} \sqcap \neg \oent{Herbivore}$ has instances at line \ref{alg:testSubClassOf:checkInconsistent}, which it does not, then named subclasses at line \ref{alg:testSubClassOf:checkIncoherent}, which it does not.
  Then it checks if it is satisfiable at line \ref{alg:testSubClassOf:checkAbsent}, which it is because the ontology does not entail that it is empty, so the algorithm returns absent.
  This means that the axiom is not entailed but it would not cause inconsistency or incoherence if added to the ontology.

  Note that it is not possible to test this axiom with any of the tools identified in section \ref{sec:related} because the left hand side of the axiom, the subclass, is not a named class.
\end{example}

% \begin{example}
%   Test if $(\oent{Frank}, \oent{Susan}) : \oent{eats}$.
%   We call algorithm \ref{alg:testObjectPropertyAssertion}
%   \[ \Call{testObjectPropertyAssertion}{\oent{eats}, \oent{Frank}, \oent{Susan}} \]
%   This first checks if $\oent{Susan}$ is one of the inferred object property values which $\oent{Frank}$ $\oent{eats}$.
%   It is not, so we enter the for-loop.
%   $N$ takes the named classes of which $\oent{Susan}$ is an inferred instance, which are $\{\oent{Giraffe}, \oent{Mammal}, \oent{Animal}\}$.
%   For each of these, we check if $\oent{Frank}$ is an instance of $\neg \exists \oent{eats} . N$.
%   When $N$ is $\oent{Animal}$, this is true, so the algorithm returns inconsistent.
% \end{example}

\begin{example}
  \label{exmp:testClassAssertion}
  Test that $\oent{Susan} : \oent{Plant}$.  Now we call algorithm \ref{alg:testClassAssertion}.
  \[ \Call{testClassAssertion}{\oent{Plant}, \oent{Susan}} \]
  This first checks if $\oent{Susan}$ is a known instance of $\oent{Plant}$ at line \ref{alg:testClassAssertion:checkEntailed}, which it is not.
  Then it checks if $\oent{Susan}$ is an instance of $\neg \oent{Plant}$ at line \ref{alg:testClassAssertion:checkInconsistent}, which it is because $\oent{Giraffe}$ is disjoint with $\oent{Plant}$, so the algorithm returns inconsistent.
\end{example}

% Alternatives:
% \[ \Call{testClassAssertion}{\exists \oent{eats}^- .\, \oent{Herbivore}, \oent{Susan}} \]
% \[ \Call{ObjectPropertyRange}{\oent{eats}, \oent{Animal}} \]

\end{document}
